---
title: "Hands-on CRM - Fitting the Model"
author: "Kristian Brock"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 10)
knitr::opts_chunk$set(fig.height = 8)
```

## Introduction
This R-notebook lets you fit the Continual Reassessment Method (CRM) to some observed DLT observations in a dose-escalation trial.

We will require several R packages. 
We should have already installed those together.
If not, run through `InstallPrerequisites.R` to check everything is present before you run through this workbook.

```{r}
library(dfcrm)
library(dplyr)
library(ggplot2)
library(trialr)
```

## Model parameters
We must define the rate of toxicity that we target:
```{r}
target <- 0.25
```

Necessarily with a CRM approach, we expect that the probabilities of efficacy and toxicity will increase with dose.
This target is the rate of toxicity that we deem justifiable in pursuit of efficacy.

We must also specify the skeleton, our initial guess at the probability of toxicity at each dose.
If we have a view on the rate of toxicity at each dose, we can define this explicitly as a vector of numerical values in R:
```{r}
skeleton <- c(0.01, 0.05, 0.1, 0.25, 0.5)
```

Alternatively, if we only have an idea at which dose the toxicity target will be, the `getprior` function in the `dfcrm` package will generate a skeleton for us.
Let us generate a skeleton with `getprior` that expects the third of five doses to be TD25:

```{r}
skeleton <- getprior(halfwidth = 0.1, target = target, nu = 3, 
                     nlevel = 5) %>% round(3)
skeleton
```

The parameters above warrant some explanation.
Higher values for the `halfwidth` parameter will give steeper curves. 
Using the value 0.05 or 0.1 normally suffices.
We tell the function our toxicity `target`, and also the position of the dose that we expect to be closest to the target using `nu`.
Lastly, we tell the function how many doses we are investigating via `nlevel`.
The `getprior` function then fits a good candidate skeleton to these parameters.
The bit of code that says `%>% round(3)` merely expresses the vector of numbers to three decimal places.

We can visualise the skeleton generated:

```{r}
num_dose <- length(skeleton)
dose_data <- data.frame(Dose = 1:num_dose, ProbTox = skeleton)

ggplot(dose_data, aes(x = Dose, y = ProbTox)) + 
  geom_point() + 
  geom_line() + 
  geom_hline(yintercept = target, col = 'red', linetype = 'dashed')
```

The function ensures that the skeleton is smoothly increasing with plausibly-spcaed out toxicity probabilities.
The plot above shows that dose 3 is our prior guess for TD25, as stipulated by `nu = 3` in the call to `getprior`.

## Fitting model to data

Let us say we have observed outcomes `1NNN 2TNT`.
Given our parameterisation, we wonder what the model will recommend next.
We must *fit* the model to the data.
This means calculating values for the free parameters in the model that best support the observed outcomes and model assumptions.

We can _parse_ this outcome string to create a data-frame reflecting doses given and toxicity outcomes:
```{r}
outcomes <- '1NNN 2TNT'
outcomes_df <- df_parse_outcomes(outcomes, as.list = FALSE)
outcomes_df
```

We can then fit the model to the outcome data:
```{r}
crm_fit1 <- crm(prior = skeleton, target = target, 
                tox = outcomes_df$tox, level = outcomes_df$doses)
```

We had to tell the `crm` function in the `dfcrm` package:

* our dose-toxicity skeleton, using the alias it expects, `prior`
* our toxicity `target`
* the observed sequence of whether DLT was observed or not using the indicative values 0 & 1, under the alias `tox`
* the administered sequence of dose indices, i.e. 1 for dose 1, etc, under the alias `level`.

Notice that the `tox` and `level` sequences were just columns plucked from the `outcomes_df` object that we created?
We could instead have written:

```{r}
crm_fit1 <- crm(prior = skeleton, target = target, 
                tox = c(0,0,0, 1,0,1), level = c(1,1,1, 2,2,2))

```

This would have yielded the same results.
In the first example, the `df_parse_outcomes` function saved us some tedious typing.

The returned object called `crm_fit1` contains information on the fit model, including the model's recommendation for the next dose:

```{r}
crm_fit1
```

The `Ptox` column shows the _posterior_ estimate of the probabilities of toxicity at each dose.
This estimate combines our initial beliefs that were reflected by the `skeleton` and the outcomes observed.
We see that the estimated toxicity rates are quite high.
This is because we have observed two DLTs in three patients at dose 2, a realised DLT-rate much higher than the 8% we specified in our `skeleton`.
The columns `LoLmt` and `UpLmt` show the approximate lower and upper confidence interval limits on the toxicity rate.
We see that those intervals are quite wide.
For example, although dose 2 is now expected to produce DLT with probability 36%, the confidence interval spans 11% to 63%, a stark demonstration of how uncertain this model still is.
After all, we have only evaluated six patients!
The model will get more precise with greater sample size.

The dose recommended by the model for the next group of patients is dose 1 because it has posterior probability of toxicity that is closes to the toxicity target.

We can visualise our posterior dose-toxicity beliefs and our skeleton on the same plot:

```{r}
dose_data <- data.frame(
  Dose = 1:num_dose, 
  ProbTox = c(crm_fit1$ptox, skeleton),
  Series = rep(c('Posterior', 'Skeleton'), each = num_dose)
)

ggplot(dose_data, aes(x = Dose, y = ProbTox)) + 
  geom_point() + 
  geom_line(aes(col = Series)) + 
  geom_hline(yintercept = target, col = 'red', linetype = 'dashed')
```

Notice how the posterior curve is always higher than the skeleton?
This is because the outcomes observed more toxicity than was expected by our skeleton.
The posterior belief is a combination of the skeleton and the data.
This means that the red dashed line representing the toxicity target intersects the posterior line at a lower dose-level, thus de-escalation is warranted.

## Exercise

Can you fit a CRM model to the outcomes `1NNN 2TNT 1NNN 2NNN`?

* What dose is recommended for the fifth cohort? 
* Why?
* Can you visualise the posterior and skeleton dose-toxicity curves?
* How close is the model to recommending the next higher dose?


