---
title: "Fixed-n trial design exercises"
author: "NAME HERE"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

This notebook will take you through the exercises in the 3.3 - Fixed-n phase II designs lecture



## Source files and packages
```{r}
source()
source()
library(clinfun)
nfleming <- function(alpha, beta, p0, p1) {
  n = (((qnorm(alpha / 2) * sqrt(p0 * (1 - p0))) + 
          (qnorm(beta / 2) * sqrt(p1 * (1 - p1))))^2) / (p1 - p0)^2
  print(n)
}
```

## Exercise - Calculate $N_F$ for given parameters
A clinician wishes to run a phase II trial to investigate a new immunotherapy treatments 
effectiveness in treating late stage Non-Small-Cell Lung cancer. A response rate lower than 
25% would indicate the treatment is ineffective, whilst a response rate higher than 35% would
warrant further investigation in a phase III trial. Determine N as per flemings design provided,
the type I and type II error rates are set at 0.10 and 0.20 respectively. 


```{r}
nfleming(alpha, beta, p0, p1)
```

## Exercise - Determine optimal and minimax designs for Simon's two-stage
For the same proposed trial above determine n for the optimal and minimax Simon's designs
```{r}
ph2simon(pu, pa, ep1, ep2, nmax)
```

Investigate what happens to n, if response rate of interest is increased by 10%
```{r}
ph2simon(pu, pa, ep1, ep2, nmax)
```

Investigate what happens to n, if you lower the type I error rate
```{r}
ph2simon(pu, pa, ep1, ep2, nmax)
```
> Do these changes to n make sense?

## Exercise - Calculate error rates for a Simon’s two-stage design
If we are instead limited by recruitment in that we can only recruit 60 patients maximum.
Calculate the error rates for a design with 20 patients in stage 1 and a stage 1 critical value of 4, 
with an additional 40 patients in stage 2 with a final critical value of 18.
```{r}
Simons2Stage(n1, r1, n, r, p0, p1, n2)
```

## Exercise - Calculate Jung’s design error rates 
Determine the type I and type II error rates for a single stage Jung's design, 
with an an overall sample size of 54, and an overall critical required difference of 5. The values of
0.15 and 0.30 are the response rates for the control and experimental arms respectively.
```{r}
Jungserr(n, a, p0, p1)
```
>Try to vary n and a, what do you notice about the error rates?





















